A fast unified model for parsing and sentence understanding

### 2016-03-15 writing/hard_stack_paper/paper.tex
Incorporate Hieu's comments.

Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences. However, they suffer from two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks. We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser. This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy. We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.

@@

TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons. Batched computation---performing synchronized computation across many examples at once---is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets (\question{What, if anything, can we cite here?}). Batching yields order-of-magnitude improvements in standard CPU implementations of neural network models, and substantial additional gains on GPUs. TreeRNNs build the parse structure of each sentence they process into the structure of the network itself, forcing the model to perform a different sequence of computations for each sentence, and thereby requiring that each sentence be processed one-by-one, making batched computation impossible in standard implementations. One aim of this paper is to restructure the computations involved in a TreeRNN in order to enable batching and make these large-scale evaluations possible.  In addition to their slow speed, TreeRNNs have an additional feature that makes them impractical in many settings: they can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications. This paper additionally aims to integrate parsing into TreeRNN processing to circumvent this issue.

@@

The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations. Below $\mid$ denotes an infix \textit{cons} function. \todo{[JG] Clarify prev. sentence.}
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
\end{description}

@@

In the SPINN, each row of the stack and each row of the buffer contains a pair $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence. \todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}

@@

The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see below). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

While it is possible to use the model that we have described thus far as a sentence encoder, at least when an external parser is available to supply \shift~and \reduce~instructions (and we do evaluate such a model), our full model includes an additional component: the tracking LSTM. (\todo{[SB] Update framing of prev sentence.}) The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which is what allows the model to stand alone as a parser, and it additionally supplies a secondary input to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation. Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function.

@@

\paragraph{Tracking left context for interpretation} Lexical ambiguity is omnipresent in natural language. Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word (first incorporates it into a partial sentence representation), the model has access to a state vector that summarizes all of the words to the left of the current word, and that state vector can serve as a crude but effective form of context to help with disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is strictly less than the sequence model has access to, and is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

>>>>

Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences. However, they suffer from two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks. We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser. This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy. We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models. \todo{[SB] Quantify success.}

@@

TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons. Batched computation---performing synchronized computation across many examples at once---is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets (\question{What, if anything, can we cite here?} \todo{[SB] Bengio 2003?}). Batching yields order-of-magnitude improvements in standard CPU implementations of neural network models, and substantial additional gains on GPUs. TreeRNNs build the parse structure of each sentence they process into the structure of the network itself, forcing the model to perform a different sequence of computations for each sentence, and thereby requiring that each sentence be processed one-by-one, making batched computation impossible in standard implementations. One aim of this paper is to restructure the computations involved in a TreeRNN in order to enable batching and make these large-scale evaluations possible.  In addition to their slow speed, TreeRNNs have an additional feature that makes them impractical in many settings: they can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications. This paper additionally aims to integrate parsing into TreeRNN processing to circumvent this issue.

@@

The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations. Below $\mid$ denotes an infix \textit{cons} function. \todo{[JG] Clarify prev. sentence. Indicate which side of the stack/buffer is the head below.}
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
\end{description}

@@

In the SPINN, each element of the stack and each element of the buffer contains a pair $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence. \todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}

@@

The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see below \todo{[SB] Ref.}). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

While it is possible to use the model that we have described thus far as a sentence encoder, at least when an external parser is available to supply \shift~and \reduce~instructions (and we do evaluate such a model), our full model includes an additional component: the tracking LSTM. (\todo{[SB] Update framing of prev sentence.}) The tracking LSTM is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition decision function, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (\todo{[SB] Refer to eq no. here and below.}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

Technically, the tracking LSTM takes three vectors as inputs at each step: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ which would be composed in a \reduce~operation. Its output hidden state at each step is used as the external input to the TreeLSTM composition function for that step and as the sole input to the transition prediction function. \todo{[SB] Trim.}

@@

\paragraph{Tracking left context for interpretation} \todo{[SB] Trim.} Lexical ambiguity is omnipresent in natural language. Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word (first incorporates it into a partial sentence representation), the model has access to a state vector that summarizes all of the words to the left of the current word, and that state vector can serve as a crude but effective form of context to help with disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is strictly less than the sequence model has access to, and is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.
